Formalizing Stack Safety as a Security Property

### 2022-09-01 submission5/onlyformalism.tex
Pretty well done

Collectively addresses and registers are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \WORDS + \REGS\).
%
A {\em machine state} \(\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\)
is a mapping from state elements to values. For convenience, we abbreviate
multiple updates \(m[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
as \(m \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
are fomulae with free variables bound to sets in \(C\).
The base machine step function
\(\mach \stepstounder{} \mach' \in \MACHS \rightarrow \MACHS\) represents a single step.

@@

\begin{align*}
\psi \in \Psi ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
| & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
| & \mathbf{return} & \\
| & \mathbf{alloc} ~ size & size \in \mathbb{N} \\
| & \mathbf{setex} \\
| & \mathbf{throwex} \\
\end{align*}

@@

Now we're ready to introduce the overlay semantics.
Additionally, we keep track of security context information
specific to each stack over the course of the run.
This data is not relevant to execution,
it is only used to define the stack safety property. We define a set of
{\it security classes}:
\[sc \in SEC ::= \public | \sealed | \unsealed\]
A {\it security view} \(V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\) divides the machine
elements in terms of their purpose relative to the active function.
The {\it initial view} \(V_0\) maps all stack locations and non-public registers
to \(\unsealed\) and all other locations and public registers to \(\public\).
A set of {\it targets} identify the addresses of the program counter or
stack pointer at a return point:
\[\begin{split}
tar \in TAR ::= & \mathit{Return} ~ a_{pc} ~ a_{sp} \\
& \mathit{Catch} ~ a_{pc} ~ a_{sp} \\
& \mathit{Done} \\
\end{split}\]
And a {\it stack context} is a non-empty stack of pairs of targets and views:
\[\context \in \CONTEXTS ::= \mathit{list} ~ (\mathit{TAR} \times \mathit{VIEW})\]

The initial stack context \(\context_0\) is \(\lbracket (\mathit{Done}, V_0) \rbracket\).

@@

@@

\judgmenttwobrlong{\((\mach,\context) \stepstounder{\psi?} (\mach',\context')\)}
                  {\(|\context'| > |\context''|\)}
                  {\(\context'' \downarrow (\mach',\context') \hookrightarrow \obsT\)}
                  {\(\context'' \downarrow (\mach,\context) \hookrightarrow \notfinished{(\mathit{observe} ~ \psi? ~ \mach}{\obsT}\)}

@@

To update the context for a call, we pair the updated view with the return target of the
call and push it to the stack. The return target will tell us whether a future
return is to the proper location with the stack pointer restored.
The view tells us how to treat the security of the callee's state, and as
it becomes more complicated, will benefit from the history that we maintain.
We take the head of a list, \(\mathit{hd}\), in the usual way, here a
total function because the context will never be empty.

On a return, we pop the topmost pair from the context that has
a \(\mathit{return}\), along with any other pairs we pass.

@@

\judgmenttwobrlong[CallC]
                  {\(\psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}}\)}
                  {\(\context' = (\mathit{Return} ~ (m[pc] + 4) ~ m[sp], V)::\context\)}
                  {\(V = \mathit{update\_view} ~ (snd ~ (hd ~ \context)) ~ m[sp] ~ \overline{r_{args}}\)}
                  {\(\mach, \context \constep{\psi} \context'\)}

@@

\caption{Call, enter, and return macro-instructions}
\label{fig:callcontexts}
\end{figure*}

@@

We start by defining a second-order logical operator
\(\context \uparrow P\), read ``\(P\) holds on return to \(\context\),''
where \(P\) is a predicate on machine states. This is a coinductive relation
similar to ``weak until'' in temporal logic -- it holds if the condition
is never reached.

\judgment{\(P ~ \mach\)}
         {\((\context \uparrow P) ~ (\mach, \pol, \prov, \context)\)}

\judgmenttwobrlong{\(\context_0 \not = \context\)}
                  {\((\context_0 \uparrow P) ~ (\mach', \pol', \prov', \context')\)}
                  {\((\mach, \pol, \prov, \context) \stepstounder{\rho,\psi?}
                    (\mach', \pol', \prov', \context')\)}
                  {\((\context_0 \uparrow P) ~ (\mach, \pol, \prov, \context)\)}

Similarly, in {\it caller confidentiality}, we will want to compare future states,
so we give a binary equivalent, \(\context \Uparrow R\), so that
\(\sigma (\context \Uparrow R) \sigma'\) holds if \(R\) holds on the
first return states of after \(\sigma\) and \(\sigma'\).

\judgment[Returned]
         {\(\mach ~ R ~ \mach'\)}
         {\((\mach,\pol,\prov,\context) ~ (\context \Uparrow R) ~ (\mach',\pol',\prov',\context)\)}

\judgmentthree[Left]
              {\(\context \not = \pi_c(\sigma_1)\)}
              {\(\sigma_1 \stepstounder{\rho;\psi?} \sigma_1'\)}
              {\(\sigma_1' ~ (\context \Uparrow R) ~ \sigma_2\)}     
              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2\)}

\judgmentthree[Right]
              {\(\context \not = \pi_c(\sigma_2)\)}
              {\(\sigma_2 \stepstounder{\rho;\psi?} \sigma_2'\)}
              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2'\)}     
              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2\)}

\paragraph*{Variants and Safe Elements}

One or more elements of a given state are {\it safe} if their values
in that state have no observable influence on future execution. We define this
formally as a sort of non-interference: for a state \(\sigma\) and
a set of elements \(\components\), \(\mathit{safe} ~ \sigma ~ \components\)
holds if we can replace the contents of \(\components\) with arbitrary values
and have the resulting state produce an equivalent trace. The state created
by modifying \(\components\) in this way is termed a variant.

@@

\definition Full states \((\mach,\pol,\prov,\context)\) and \((\nach,\pol,\prov,\context)\)
are also {\em \(\components\)-variants} if \(\mach\) and \(\nach\) are.

\definition An element set \(\components\) is safe in state \(\sigma)\),
written \(\components ~ \mathit{SAFE} ~ \sigma\), if for all
\(\sigma'\) such that \(\sigma \approx_{\components'} \sigma'\), if 
\(\sigma \hookrightarrow \obsT\) and
\(\sigma' \hookrightarrow \obsT'\), then
\(\obsT \eqsim \obsT'\).

@@

\definition Let \(\sigma = (\mach,\pol,\prov,\context)\) and
\(\components = \{\component | \mathit{hd\_def} ~ \context ~ \component = \sealed\}\).
Let \(P\) be a predicate on states \(\sigma'\) that holds if
\(\Delta(\mach, \pi_m(\sigma')) \cap \components ~ \mathit{SAFE} ~ \sigma'\).
Then \(\sigma\) enjoys {\it return-time integrity} if \(\context \downarrow P\) holds.

\definition A system enjoys {\it stack integrity} if, for all reachable states \(\sigma\) such that
\(\sigma \stepstounder{\rho;\mathbf{Call}\dots} \sigma'\),
\(\sigma'\) enjoys return-time integrity.

@@

We could define a natural, strict form of confidentiality that requires that the final
states be identical. However, as with integrity, this is stricter than we need -- there
may be some conditions in which a secret is copied, but never used again. So, we once
again wish for our final condition to make the values {\it safe}. Which values?

@@

\definition Let \(\sigma = (\mach,\pol,\prov,\context)\) be a state and
\(\components = \{\component | \mathit{hd\_def} ~ \context ~ \component \in
\{\sealed, \unsealed\}\}\).

Then \(\sigma\) enjoys {\it caller confidentiality} if, for any \(\sigma'\)
that is a \(\components\)-variant of \(\sigma\), we can take
\(\context \uparrow \sigma \hookrightarrow \obsT\) and
\(\context \uparrow \sigma' \hookrightarrow \obsT'\) and have that
\(\obsT \simeq \obsT'\). And if, additionally, \(\sigma ~ (\context \Uparrow R) ~ \sigma'\),
where \(\sigma_1 ~ R ~ \sigma_2\) if
\(\bar{\Diamond}(\mach,\pi_m(\sigma_1),\pi_m(\sigma'),\pi_m(\sigma_2)) ~ \mathit{SAFE} ~ \sigma_1\).

\definition A system enjoys {\it universal caller confidentiality} if, for all reachable states
\(\sigma\) such that \(\sigma \stepstounder{\rho;\mathbf{Call}\dots} \sigma'\),
\(\sigma'\) enjoys caller confidentiality.

\subsection{The Machine in Detail}
\label{sec:prelim}

The building blocks of the machine are {\em words} and {\em registers}.
Words (\(\WORDS\)) are 64-bit integers ranged over by \(w\) and, when used as addresses,
\(a\). Each processor has 32 registers, ranged over in general by \(r\),
plus a program counter (\(\PCname\)). The following table divides
the registers into metavariables, some belonging to caller- or callee- sets:

We specify that a register belongs to processor \(\rho \in \PROCS\) by writing
it \(r^\rho\), but will omit this notation when the active
processor is unambiguous from context. The set of all registers across all processors
is \(\REGS\).

Collectively addresses and registers are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \WORDS + \REGS\).
%
A {\em machine state} \(\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\)
is a mapping from state elements to values. For convenience, we abbreviate
multiple updates \(m[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
as \(m \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
are fomulae with free variables bound to sets in \(C\).
The base machine step relation 
\(\mach \machstep{\rho} \mach' \subseteq \MACHS \times \PROCS \times \MACHS\)
represents a single step of the processor \(\rho\).

The machine is equipped with additional state representing the tag-based hardware.
We represent this as a policy state, \(\pol \in \POLS ::= \COMPONENTS \rightarrow \mathcal{T}\),
where \(\mathcal{T}\) is the set of tags, discussed [in the section on security.]
We represent a {\it policy} via the step relation
\(\mach, \pol \harpoonunder{\rho}_P \pol' \subseteq \MACHS \times \POLS \times \PROCS \times \POLS\).
Steps are deterministic and  undefined represents a {\it failstop}.

Our properties additionally depend on a rudimentary model of memory-safe
objects and pointer provenance. A provenance \(\phi \in \mathbb{N}\)
identifies a value derived from the \(\phi\)th object allocated since
the beginning of execution. A provenance map
\(\prov \in \PROVS ::= \COMPONENTS \rightarrow 2^{\mathbb{N}}\)
designates, for every machine element, which provenances its current
value contains. We define a step relation on provenance maps
\(\mach, \prov \harpoonunder{\rho}_{pm} \prov' \subseteq \MACHS \times
\PROVS \times \PROCS \times \PROVS\),
exerpted in Figure \ref{fig:provenance} [TODO].

We define a step relation over triples of a machine state,
a policy state, and a provenance map:

\judgmentthree{\(\mach \machstep{\rho} \mach'\)}
              {\(\mach, \pol \polstep{\rho} \pol'\)}
              {\(\mach, \prov \provstep{\rho} \prov'\)}
              {\((\mach,\pol,\prov) \stepstounder{\rho} (\mach',\pol',\prov')\)}

As the last piece of state, we keep track of security context information
specific to each stack
over the course of the run. This data is not relevant to execution,
it is only used to define the stack safety property. We define a set of
{\it security classes}:
\[sc \in SEC ::= \public | \sealed | \mathit{object}(\phi) | \unsealed\]
And of {\it targets}, which identify the addresses of the program counter or
stack pointer at a return or catch point:
\[\begin{split}
tar \in TAR ::= & \mathit{Return} ~ a_{pc} ~ a_{sp} \\
| & \mathit{Catch} ~ a_{pc} ~ a_{sp} \\
\end{split}\]
A {\it security view} \(V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\) divides the machine
elements in terms of their purpose relative to the active function.
And a {\it stack context} is a stack of pairs of targets and views:
\[\context \in \CONTEXTS ::= \mathit{list} ~ (\mathit{TAR} \times \mathit{VIEW})\]

The set of instructions, \(I \subseteq \WORDS\), corresponds to the RISC-V ISA.
We further introduce {\it macro-instructions}: sequences of
instructions that operate as if they were single, atomic pseudoinstructions.
A macro-instruction consists of both a list
of instructions and a pseudoinstruction that identifies its purpose in the
property. The pseudoinstructions, and the data that they carry, are as follows.
[TODO: still handle stack args by offset.]

\begin{align*}
\psi \in \Psi ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
| & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
| & \mathbf{return} & \\
| & \mathbf{alloc} ~ size & size \in \mathbb{N} \\
| & \mathbf{setex} \\
| & \mathbf{throwex} \\
\end{align*}

\judgment[Done]
         {\(\mach[\PCname] < a_{low} \lor \mach[\PCname] > a_{high}\)}
         {\(a_{low},a_{high} \vdash (\mach,\pol,\prov) \manystepstounder{\rho} (\mach,\pol,\prov)\)}

\judgmenttwobrlong[Step]
                  {\(a_{low} \leq \mach[\PCname] < a_{high}\)}
                  {\((\mach,\pol,\prov) \stepstounder{\rho} (\mach',\pol',\prov')\)}
                  {\(a_{low},a_{high} \vdash (\mach',\pol',\prov') \manystepstounder{\rho} (\mach'',\pol'',\prov'')\)}
                  {\(a_{low},a_{high} \vdash (\mach,\pol,\prov) \manystepstounder{\rho} (\mach'',\pol'',\prov'')\)}

Now we're finally ready to introduce the overlay semantics.
A {\it full state} \(\sigma\) consists of a 4-tuple of a machine state,
a policy state, a provenance map, and a map from processors to stack contexts:
\[\sigma = (\mach, \pol, \prov, ctxs) \in \mathcal{S} ::=
\MACHS \times \POLS \times \PROVS \times (\PROCS \rightarrow \CONTEXTS)\]

The execution relation for full states
\(\sigma \stepstounder{\rho;\psi?} \sigma' \subseteq
\mathcal{S} \times \PROCS \times (\Psi + \bot) \times \mathcal{S}\)
proceeds by an {\it overlay step}
if the program counter points to the start of a macro-instruction.
Otherwise, it proceeds by the default step. The definitions of context steps
\(\constep{\rho;\psi}\) will be defined for each pseudoinstruction as we introduce
topics.

\judgmenttwobrlongbrlong[Overlay]
                        {\((\bar{i}, \psi) \in J\)}
                        {\(\mach',\prov',(ctxs ~ \rho) \constep{\rho;\psi} \context\)}
                        {\((\mach,\pol,\prov) \manystepstounder{\rho} (\mach',\pol',\prov')\)}
                        {\(\mach[\PCname+n] = \bar{i}[n]\) for each \(n\) where \(0 \leq n < |\bar{i}|\)}
                        {\((\mach,\pol,\prov,ctxs) \stepstounder{\rho;\psi} (\mach',\pol',\prov',ctxs[\rho \mapsto \context])\)}

\judgment[Default]
         {\((\mach,\pol,\prov) \stepstounder{\rho} (\mach',\pol',\prov')\)}
         {\((\mach,\pol,\prov,ctxs) \stepstounder{\rho; \bot} (\mach',\pol',\prov',ctxs)\)}

\paragraph*{Concurrency and Traces}

So far we have defined the execution steps of single processors. Processors
have their own register banks, but share memory, and may step independently
in any order. We can represent sequences of these steps as traces.
Some steps of execution can be interpreted as {\it events} that can be observed
from outside the system. We define events \(\obs \in \OBSS\) as calls, labeled
with the target address and the values of the arguments, or else as the silent
event \(\tau\):
\[\begin{split}
\obs \in \OBSS ::= & ~ \mathbf{callE} ~ a_{target} ~ \overline{w_{args}} \\
| & ~ \tau \\
\end{split}\]

\begin{figure*}
\[\mathit{observe} ~ \rho ~ \psi? ~ m =
\begin{cases}
  (\rho, \mathbf{callE} ~ m[\PCname] ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket)
  & \textnormal{when } ~ \psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}} \\
%  (\rho, \mathbf{returnE} ~ m[\PCname] & \textnormal{when } \psi = \mathbf{return} \\
  (\rho, \tau) & \textnormal{otherwise} \\
\end{cases}\]
\caption{Converting overlay steps to events}
\label{fig:observe}
\end{figure*}

\judgment{\(\not \exists \sigma' . \sigma \stepstounder{} \sigma'\)}
         {\(\sigma \hookrightarrow \tau\)}

\judgmenttwo{\(\sigma \stepstounder{\rho; \psi?} \sigma'\)}
            {\(\sigma' \hookrightarrow \obsT\)}
            {\(\sigma \hookrightarrow \notfinished{(\rho,\mathit{observe} ~ \rho ~ \psi? ~ \pi_m(\sigma'))}{\obsT}\)}

We define another relation that takes a trace of a callee up to its return, if one occurs.
In this case we must specify that we are considering a single processor.
We write this \(\rho,\context \downarrow \sigma \hookrightarrow \obsT\), where \(\context\)
is the context of the caller.

\judgment{\(\context = \pi_c(\sigma) ~ \rho\)}
         {\(\rho, \context \downarrow \sigma \hookrightarrow \tau\)}

\judgmentthree{\(\sigma \stepstounder{\rho',\psi?} \sigma'\)}
              {\(\rho \not = \rho'\)}
              {\(\rho, \context \downarrow \sigma' \hookrightarrow \obsT\)}
              {\(\rho, \context \downarrow \sigma \hookrightarrow \obsT\)}

\judgmentthree{\(\sigma \stepstounder{\rho,\psi?} \sigma'\)}
              {\(\context \not = \pi_c(\sigma') ~ \rho\)}
              {\(\context \downarrow \sigma' \hookrightarrow \obsT\)}
              {\(\context \downarrow \sigma \hookrightarrow \notfinished{(\mathit{observe} ~ \rho ~ \psi? (\pi_m{\sigma'}))}{\obsT}\)}

We say that two event traces $\obsT_1$ and $\obsT_2$ are {\em similar}
from the perspective of processor \(\rho\), written \(\obsT_1 \eqsim_\rho
\obsT_2\), if the sequence of non-silent \(\rho\) events is the same. That is, we
compare up to deletion of \(\tau\) events and events from other processors, coinductively:

\begin{minipage}{.4\columnwidth}
  \judgment{}{\(\obsT \eqsim_\rho \obsT\)}
\end{minipage}
\begin{minipage}{.4\columnwidth}
  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
           {\(\notfinished{(\rho, \obs)}{\obsT_1} \eqsim_\rho \notfinished{(\rho, \obs)}{\obsT_2}\)}
\end{minipage}

\begin{minipage}{.4\columnwidth}
  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
           {\(\notfinished{(\rho,\tau)}{\obsT_1} \eqsim_\rho \obsT_2\)}
\end{minipage}
\begin{minipage}{.4\columnwidth}
  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
           {\(\obsT_1 \eqsim_\rho \notfinished{(\rho,\tau)}{\obsT_2}\)}
\end{minipage}

\begin{minipage}{.4\columnwidth}
  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
           {\(\notfinished{(\rho',\obs)}{\obsT_1} \eqsim_\rho \obsT_2\)}
\end{minipage}
\begin{minipage}{.4\columnwidth}
  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
           {\(\obsT_1 \eqsim_\rho \notfinished{(\rho',\obs)}{\obsT_2}\)}
\end{minipage}

In this section we will present the context steps for allocations,
calls, and returns, and show how they
translate into properties. They are given formally in Figure \ref{fig:callcontexts}.
Allocations are simple: from the current stack pointer, a range up to the size of
the allocation is marked \(\mathit{object}(\phi)\) for a fresh identifier \(\phi\).
[TODO: actually generate fresh \(\phi\) by keeping counter...]
[TODO: need to put the original \(\phi\) somewhere. Maybe macro-instruction
  for offsetting from stack pointer? Ugh.]

Alloc macro-instructions model allocations that might, or might not, be
shared. A function can always allocate more opjects by adjusting its
stack pointer outside of a macro-instruction, but this will not be reflected
in the context.

First, we define a function, {\it update view}, that defines the view of a callee
in terms of that of the callee, together with the provenance map.
First, all stack locations below the stack pointer
are \(\unsealed\) -- fair game for the callee to allocate or write to, and
to read after writing. Above the stack pointer, anything that is not an object
becomes \(\sealed\). Argument and system registers are automatically marked
\(\mathit{public}\), and all others are \(\sealed\).
Finally, we must address stack objects: we recursively identify all objects
that are reachable via pointer provenances found in \(\mathit{public}\)
registers and in reachable objects, and these retain their view, while
all others become \(\sealed\). [TODO: implement \(\mathit{reach}\) to do this.]

To update the context, we pair the updated view with the return target of the
call: the return target tells us whether a return is to the proper location
which the stack pointer restored, and the view tells us how to treat the
security of the callee's state. We define \(\mathit{hd\_def}\) as taking
a context and returning its head, or \(((0,0),\lambda \component.\unsealed)\)
if the context is empty. [TODO: structure it so that we don't need to return
  a dummy target.]

>>>>

Collectively addresses and registers are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \WORDS + \REGS\).
%
A {\em machine state} \(\mach \in \MACHS\) is a type that exposes a mapping
\(\mach[\component]\) from state elements to values, and may also contain additional
security state. For convenience, we abbreviate
multiple updates \(m[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
as \(m \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
are fomulae with free variables bound to sets in \(C\).
The base machine step function
\(\mach \stepstounder{} \mach' \in \MACHS \rightharpoonup \MACHS\) represents a single step.
It is partial to represent the capacity of the machine to failstop.
In order to accommodate failstop behavior, our properties will be {\it termination insensitive}.

@@

\begin{align*}
\psi \in \Psi ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
| & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
| & \mathbf{return} & \\
| & \mathbf{alloc} ~ size & size \in \mathbb{N} \\
| & \mathbf{setex} ~ \addr_{buf} \\
| & \mathbf{throwex} ~ \addr_{buf} \\
\end{align*}

@@

Now we're ready to introduce the overlay semantics.
Additionally, we keep track of security context information
specific to each stack over the course of the run.
This data is not relevant to execution,
it is only used to define the stack safety property. We define a set of
{\it security classes}:
\[sc \in SEC ::= \public | \sealed | \unsealed\]
A {\it security view} \(V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\) divides the machine
elements in terms of their purpose relative to the active function.
The {\it initial view} \(V_0\) maps all stack locations and non-public registers
to \(\unsealed\) and all other locations and public registers to \(\public\).
A set of {\it targets} identify the addresses of the program counter or
stack pointer at a return point:
\[\begin{split}
tar \in TAR ::= & \mathit{Return} ~ a_{pc} ~ a_{sp} \\
& \mathit{Catch} ~ a_{buf} ~ a_{pc} ~ a_{sp} \\
& \mathit{Done} \\
\end{split}\]
And a {\it stack context} is a non-empty stack of pairs of targets and views:
\[\context \in \CONTEXTS ::= \mathit{list} ~ (\mathit{TAR} \times \mathit{VIEW})\]

The initial stack context \(\context_0\) is \(\left[ (\mathit{Done}, V_0) \right]\).

@@

\judgment{\(\not \exists \mach' ~ \context'. (\mach,\context) \stepstounder{\psi?} (\mach',\context')\)}
         {\((\mach,\context) \hookrightarrow \tau\)}

@@

\judgmenttwobrlong{\((\mach,\context) \stepstounder{\psi?} (\mach',\context')\)}
                  {\(|\context'| > |\context''|\)}
                  {\(\context'' \downarrow (\mach',\context') \hookrightarrow \obsT\)}
                  {\(\context'' \downarrow (\mach,\context) \hookrightarrow \notfinished{\mathit{observe} ~ \psi? ~ \mach}{\obsT}\)}

@@

To update the context for a call, we pair the updated view with the return target of the
call and push it to the stack. The return target will tell us whether a future
return is to the proper location with the stack pointer restored.
The view tells us how to treat the security of the callee's state, and as
it becomes more complicated, will benefit from the history that we maintain.

On a return, we pop the topmost pair from the context that has
a \(\mathit{return}\), along with any other pairs (catch points,
for instance) that we pass.

@@

\judgmenttwobrlong[CallC]
                  {\(\psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}}\)}
                  {\(T' = \mathit{Return} ~ (m[pc] + 4) ~ m[sp]\)}
                  {\(V' = \mathit{update\_view} ~ V ~ m[sp] ~ \overline{r_{args}}\)}
                  {\(\mach, (T,V)::\context \constep{\psi} (T',V')::\context\)}

@@

\caption{Call and return context updates}
\label{fig:callcontexts}
\end{figure*}

@@

We start by defining a second-order logical operator
\(\context_{ret} \uparrow P\), read ``\(P\) holds on return to \(\context_{ret}\),''
where \(P\) is a predicate on machine states. This is a coinductive relation
similar to ``weak until'' in temporal logic -- it holds if the condition
is never reached.

\judgmenttwo[Returned]
            {\(P ~ (\mach,\context)\)}
            {\(|\context| \leq |\context_{ret}|\)}
            {\((\context_{ret} \uparrow P) ~ (\mach, \context)\)}

\judgmenttwobrlong[Step]
                  {\(|\context| > |\context_{ret}|\)}
                  {\((\mach, \context) \stepstounder{\psi?} (\mach', \context')\)}
                  {\((\context_{ret} \uparrow P) ~ (\mach', \context')\)}
                  {\((\context_{ret} \uparrow P) ~ (\mach, \context)\)}

Similarly, in {\it caller confidentiality}, we will want to compare future states,
so we give a binary equivalent, \(\context \Uparrow R\), so that
\((\mach,\context) ~ (\context'' \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
first return states after \((\mach,\context)\) and \((\mach',\context')\).

\judgmentthree[Returned]
              {\((\mach,\context) ~ R ~ (\mach',\context')\)}
              {\(|\context| \leq |\context_{ret}|\)}
              {\(|\context'| \leq |\context_{ret}|\)}
              {\((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\mach',\context')\)}

\judgmenttwobr[Left]
              {\(\context_1 > \context_{ret}\)}
              {\((\mach_1,\context_1) \stepstounder{\psi?} (\mach_1',\context_1')\)}
              {\(\context_2 > \context_{ret}\)}
              {\((\mach_1',\context_1') ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}
              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}

\judgmenttwobr[Right]
              {\(\context_1 > \context_{ret}\)}
              {\((\mach_2,\context_2) \stepstounder{\psi?} (\mach_2',\context_2')\)}
              {\(\context_2 > \context_{ret}\)}
              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2',\context_2')\)}
              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}

\paragraph*{Variants and Vestigial Elements}

One or more elements of a given state are {\it vestigial} if their values
in that state have no observable influence on future execution. We define this
formally as a sort of non-interference: for a state \((\mach,\context)\) and
a set of elements \(\components\), \((\mach,\context) \parallel \components\)
holds if we can replace the contents of \(\components\) with arbitrary values
and have the resulting state produce an equivalent trace. The state created
by modifying \(\components\) in this way is termed a variant.

@@

\definition An element set \(\components\) is vestigial in state \(\sigma)\),
written \((\mach,\context) \parallel \components\), if for all
\(\nach\) such that \(\mach \approx_{\components'} \mach'\), if 
\((\mach,\context) \hookrightarrow \obsT\) and
\((\nach,\context) \hookrightarrow \obsT'\), then
\(\obsT \eqsim \obsT'\).

@@

\definition Let \((\mach,(T,V)::\context)\) be a state and
\(\components = \{\component | V ~ \component = \sealed\}\).
Let \(P\) be a predicate that holds on a state \((\mach',\context')\) if
\(\mach' ~ \parallel ~ \Delta(\mach,\mach') \cap \components\).
Then \(\mach\) enjoys {\it return-time integrity} if \(\context \downarrow P\) holds.

\definition A system enjoys {\it universal stack integrity} if, for all reachable states
\((\mach,\context)\) such that
\((\mach,\context) \stepstounder{\psi} (\mach',\context')\) with \(\psi = \mathbf{Call} ~ \_ ~ \_\),
\((\mach',\context')\) enjoys return-time integrity.

@@

We could define a natural, strict form of confidentiality that requires that the final
states be identical. However, as with integrity, this is stricter than we need -- there
may be some conditions in which a secret is copied, but never used again. So, we once
again wish for our final condition to allow values to be vestigial. Which values?

@@

\definition Let \((\mach,(T,V)::\context)\) be a state and
\(\components = \{\component | V ~ \component \in \{\sealed, \unsealed\}\}\).

Then \((\mach,\context)\) enjoys {\it caller confidentiality} with respect to
some \(\context_{ret}\) if, for any \(\nach\)
that is a \(\components\)-variant of \(\mach\), we can take
\(\context \uparrow (\mach,\context) \hookrightarrow \obsT\) and
\(\context \uparrow (\nach,\context) \hookrightarrow \obsT'\) and have that
\(\obsT \simeq \obsT'\);
and additionally if \((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\nach,\context)\),
where \((\mach_1,\context_1) ~ R ~ (\nach_2,\context_2)\) if
\((\mach_1 \parallel \bar{\Diamond}(\mach,\mach_1,\nach,\nach_2)\).

\definition A system enjoys {\it universal caller confidentiality} if, for all reachable states
\((\mach,\context)\) such that \((\mach,\context) \stepstounder{\psi} (\mach',\context')\)
with \(\psi = \mathbf{Call} ~ \_ ~ \_\),
\((\mach',\context')\) enjoys caller confidentiality with respect to \(\context\).

\section{Allocating Memory}

We now introduce the context step for the \(\mathbf{alloc} ~ \mathit{size}\)
macro-instruction. Without \(\mathbf{alloc}\), our system already allowed a function
to allocate as much private memory as it liked by moving the stack pointer. With
\(\mathbf{alloc}\) it can now also allocate public objects that can be accessed
freely by its callees. Later we will refine this system to restrict public objects
to be accessed only via capabilities.

\judgmentbr[AllocC]
           {\(\psi = \mathbf{alloc} ~ size\)}
           {\(V' = V\llbracket m[sp]+i \mapsto \public | 0 < i \leq size \rrbracket\)}
           {\((\mach, (T,V)::\context) \constep{\psi} (T,V')::\context\)}

The remainder of our property definitions are unchanged. The \(\mathit{update\_view}\)
function preserves the public status of allocated objects, and they are exempt from
integrity and confidentiality properties.

\section{Exceptions}

We implement exceptions via another pair of macro-instructions:
\(\mathbf{setex} ~ a_{buf}\), akin to \(\mathit{setjmp}\), with the
jump buffer stored at \(a_{buf}\), and \(\mathbf{throwex} ~ a_{buf}\),
akin to \(\mathit{longjmp}\). The context updates for exceptions are given
in Figure \ref{fig:excontexts}. The update for \(\mathbf{setex}\) pushes
a \(\mathit{Catch}\) target onto the stack, identified by \(\addr_{buf}\),
with the same view as the current function. For \(\mathbf{throwex}\),
we pop the stack until we find a \(\mathit{Catch} ~ a_{buf} ~ \_ ~ \_\),
popping it as well. Rules apply in order.

\begin{figure*}

\judgmenttwo[SetExC]
            {\(\psi = \mathbf{setex} ~ \addr_{buf}\)}
            {\(T' = \mathit{Catch} ~ \addr_{buf} ~ (\mach[pc] + 4) ~ \mach[sp]\)}
            {\(\mach, (T,V)::\context \constep{\psi} (T',V)::(T,V)::\context\)}

\judgment[ThrowExCCatch]
         {\(\psi = \mathbf{throwex} ~ \addr_{buf}\)}
         {\(\mach, (\mathit{Catch} \addr_{buf} ~ a_{pc} ~ a_{sp}, V)::\context \constep{\psi} \context\)}

\judgment[ThrowExCDone]
         {\(\psi = \mathbf{throwex} ~ \addr_{buf}\)}
         {\(m, [\mathit{Done},V] \constep{\psi} [\mathit{Done},V]\)}

\judgmenttwo[ThrowExCSkip]
         {\(\psi = \mathbf{throwex} ~ \addr_{buf}\)}
         {\(\mach, \context \constep{\psi} \context'\)}
         {\(\mach, (\_::\context) \constep{\psi} \context'\)}

\caption{Exception context updates}
\label{fig:excontexts}
\end{figure*}

\section{Tail Call Elimination}

I submit that tail calls will work perfectly fine with all of our definitions
using the context update in Figure \ref{fig:tailcallcontexts}. All of the
``until return'' stuff works fine, because all the calls return at once.
We need to define callee confidentiality still.

\begin{figure*}

\judgmenttwo[TailCallC]
            {\(\psi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}}\)}
            {\(V' = \mathit{update\_view} ~ V ~ m[sp] ~ \overline{r_{args}}\)}
            {\(\mach, (T,V)::\context \constep{\psi} (T,V')::\context\)}

\caption{Tailcall context updates}
\label{fig:tailcallcontexts}
\end{figure*}

\section{Provenance, Capabilities, and Protecting Objects}

So far, we have given functions the ability to allocate completely unprotected
objects. What if we want to express a finer-grained notion of safety, in which
such objects are protected unless the function that owns them intentionally
passes a pointer to them?

In order to express such a property, we need our machine to carry some notion
of {\it pointer provenance} -- a distinction between a pointer that is intended to
point to a given object, and non-pointer integers as well as pointers to other objects.

\section{Simple Concurrency}
